---
title: "Simple LFM Graphs"
author: "Josie Lesage, updated by Caroline Shepherd"
date: "11/14/2023"
output: html_document
editor_options: 
  chunk_output_type: console
    theme: cerulean
    code_folding: hide
    highlight: monochrome
    toc: yes
    toc_depth: 5
    toc_float: yes
    collapsed: yes
    smooth_scroll: yes
    number_sections: no
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotrix)
library(reactable)
library(gridExtra)
library(lubridate)
library(here)
library(janitor)

```

# Introduction

This code is meant to produce a simple graph of LFM at sites near the Santa Barbara Botanic Garden to raise awareness of fire danger and keep local residents informed.  

Data is typically collected every two weeks from each site. Samples are collected and dried in accordance with USFS recommendations.  

In the near future, we hope to show conditions near the Garden (in Mission Canyon) and near St. Mary's. 

```{r data, include=FALSE, echo=FALSE}
#this section of code reads in the data and cleans it up

lfm_raw <- read.csv(here("LFM_PaintedCave_data.csv")) %>%
  clean_names() %>% 
  mutate(lfm = as.numeric(lfm)) %>% 
  mutate(lfm = lfm*100,
         species = recode_factor(species,
                                 "ADFA" = "Chamise",
                                 "CEME" = "Bigpod ceanothus")) %>%
  mutate(collection_date2 = ymd(collection_date)) %>%
  mutate(recode_factor(subsite, "Unknown" = ""))

lfm <- lfm_raw %>%
  filter(concern != "Yes",
         species != "QUAG") %>% # removing any entries that are marked with "concern" and oaks - will analyze oaks later
  group_by(collection_date, site, species, subsite) %>%
  summarise(meanLFM = mean(lfm), # summary statistics for mean and std deviation
            seLFM = std.error(lfm)) %>%
  mutate(site2 = site,
         species2 = species) %>%
  unite(site2, species2, col = "ID", sep = " - ") # creating new column with ID

```

# Current conditions 
Here are the current conditions at Painted Cave. At some sites, only one sample is collected and there are no error bars. For sites with error bars, these represent the standard error.

```{r Graph of 2021 conditions, include = FALSE}
LFM_2020 <- lfm %>%
  separate(collection_date, sep = "-", into = c("year", "month", "day")) %>%
  filter(year == "2020" | year == "2021",
         site == "Painted Cave") %>%
  unite(collection_date, "year", "month", "day", sep = "-") 
        
ggplot(LFM_2020, aes(x = collection_date, y = meanLFM, color = subsite, group = (subsite))) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM, width = 0.2), size = 1) +
  geom_line(size = 1) +
  labs(x = "Collection date",
       y = "Live Fuel Mositure Content (%)",
       title = "Live Fuel Moisture Conditions") +
  scale_y_continuous(limits = c(0,130), expand = c(0,0), breaks=seq (0,200,10)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.25, hjust = 0.25)) +
  facet_wrap(~species)

reactable(
LFM_summary <- LFM_2020 %>%
  group_by(collection_date, species) %>%
  summarise(stderr = std.error(meanLFM),
            meanLFM = mean(meanLFM)) %>%
  mutate_if(is.numeric, round, digits = 2) %>%
  mutate(lfm = str_c(meanLFM, stderr, sep = " ± ")) %>%
  select(collection_date, lfm, species) %>%
  spread(collection_date, lfm)
)

```


```{r Graph of current conditions, include = TRUE}

# recreating above code for 2022-2023

LFM_2023 <- lfm %>%
  separate(collection_date, sep = "-", into = c("year", "month", "day")) %>%
  filter(year == "2023" | year == "2022",
         site == "Painted Cave") %>%
  unite(collection_date, "year", "month", "day", sep = "-")
        
lfm2023_big <- ggplot(LFM_2023, aes(x = collection_date, y = meanLFM, color = subsite, group = (subsite))) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM, width = 0.2), size = 1) +
  geom_line(size = 1) +
  labs(x = "Collection Date",
       y = "Live Fuel Mositure Content (%)",
       title = "Live Fuel Moisture Conditions") +
  scale_y_continuous(limits = c(0,190), expand = c(0,0), breaks=seq (0,200,10)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.25, hjust = 0.25)) +
  facet_wrap(~species)

reactable(
LFM_summary <- LFM_2023 %>%
  group_by(collection_date, species) %>%
  summarise(stderr = std.error(meanLFM),
            meanLFM = mean(meanLFM)) %>%
  mutate_if(is.numeric, round, digits = 2) %>%
  mutate(lfm = str_c(meanLFM, stderr, sep = " ± ")) %>%
  select(collection_date, lfm, species) %>%
  spread(collection_date, lfm)
)

ggsave(plot = lfm2023_big, filename = "Figures/lfm2023_big.png", height = 8.5, width = 11, units = "in", dpi = 300)

lfm2023_big # showing graph in html by calling it here

```

# Historical data
We'd also like to know how current conditions compare with past data, we we'll summarize average monthly LFM for the past few years, and then will compare that line to this year's data.

```{r historical comparison}
# comparing pre-2022 avg to 2022
LFM_pre_2021 <- lfm_raw %>%
  filter(concern != "Yes", 
         species != "CEBE",
         species != "QUAG", 
         site == "Painted Cave") %>%
  separate(collection_date, sep = "-", into = c("year", "month", "day")) %>%
  filter(year <= "2021") %>%
  group_by(site, subsite, species, month) %>%
  summarise(meanLFM = mean(lfm),
            seLFM = std.error(lfm)) %>%
  mutate(time = "Historical Mean (2013-2021)")

LFM_2022 <- lfm_raw %>%
  filter(concern != "Yes",
         species != "CEBE",
         species != "QUAG", 
         site == "Painted Cave") %>%
  separate(collection_date, sep = "-", into = c("year", "month", "day")) %>%
  filter(year == "2022"| year == "2023") %>%
  group_by(site, subsite, species, month) %>%
  summarise(meanLFM = mean(lfm),
            seLFM = std.error(lfm)) %>%
  mutate(time = "2022")

LFM_comp <- bind_rows(LFM_pre_2021, LFM_2022) %>%
  ungroup() %>%
  mutate(subsite = fct_recode(subsite, "PC - Laurel Spring" = "PC - LS"),
         site2 = site,
         subsite2 = subsite,
         species2 = species,
         time2 = time) %>%
  unite(subsite2, time2, col = "ID", sep = " ")
  

# THE PROBLEM IS HERE
LFM_comp_graph <- ggplot() +
  geom_line(LFM_comp, aes(x = month, y = meanLFM, color = ID, group = ID)) +
  geom_line(LFM_2022, aes(x = month, y = meanLFM, color = subsite)) +
  geom_point(aes(color = ID)) +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM, width = 0.2), size = 1) +
  geom_line(size = 1, aes(group = ID, color = ID)) +
  labs(x = "Collection month",
       y = "Live Fuel Mositure Content (%)",
       title = "Live Fuel Moisture Conditions at Painted Cave") +
  scale_y_continuous(limits = c(45,140), expand = c(0,0), breaks=seq (0,200,10)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, vjust = -0.5, hjust = 0.5)) +
  facet_wrap(~species)

LFM_comp_graph  

## comparing all years individually 
LFM_allyears <- lfm_raw %>%
  filter(concern != "Yes",
         species != "CEBE",
         species != "QUAG", 
         species != "CECRA",
         site != "Montecito Fire",
         site != "Tunnel Road",
         site != "St Mary's") %>%
  separate(collection_date, sep = "-", into = c("year", "month", "day")) %>%
  group_by(species, year, month, subsite) %>%
  summarise(meanLFM = mean(lfm),
            seLFM = std.error(lfm)) 

LFM_all_graph <- ggplot(LFM_allyears, aes(x = month, y = meanLFM, color = year, group = year)) +
  geom_point() +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM, width = 0.2), size = 1) +
  geom_line(size = 1, aes(color = year)) +
  labs(x = "Collection date",
       y = "Live Fuel Mositure Content (%)",
       title = "Live Fuel Moisture Conditions at Painted Cave") +
  scale_y_continuous(limits = c(45,160), expand = c(0,0), breaks=seq (0,200,10)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, vjust = -0.5, hjust = 0.5)) +
  facet_wrap(~species)

LFM_all_graph

# grid.arrange(LFM_comp_graph, LFM_all_graph)

```

## For website pub.
```{r historical web}
LFM_hist <- lfm_raw %>%
  filter(concern != "Yes",
         species != "CEBE",
         species != "QUAG",
         site != "Montecito Fire",
         site != "ECC",
         site != "Tunnel Road") %>%
  mutate(collection_date2 = collection_date) %>%
  separate(collection_date, sep = "-", into = c("year", "month", "day")) %>%
  mutate(month1 = month(ymd(collection_date2), label = TRUE, abbr = FALSE)) %>%
  group_by(site, species, month1) %>%
  summarise(meanLFM = mean(lfm, na.rm=TRUE),
            seLFM = std.error(lfm)) %>%
  unite(site, species, col = "ID", sep = " - ") %>%
  mutate(ID = as.factor(ID), 
    ID = fct_relevel(ID, "Painted Cave - Chamise", 
                           "St Mary's - Chamise", 
                           "Painted Cave - Bigpod ceanothus",
                           "St Mary's - Bigpod ceanothus")) %>%
  ungroup()
# above plotted data from 2013-present for painted cave, but limited for St Mary's

LFMHist_big <- ggplot(LFM_hist, aes(x = month1, y = meanLFM, color = ID)) +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM, width = 0.2), size = 1) +
  geom_line(aes(x = month1, y = meanLFM, group = ID), size = 1.25) +
  geom_hline(aes(yintercept = 60, color = "60% Critical Threshold"), linetype="dashed", size=1) +
  labs(x = "",
       y = "Live Fuel Mositure Content (%)",
       color = "Site and Species",
       title = "Average historical fuel moisture content (per month & species) \n at each collection site") +
  scale_y_continuous(limits = c(45,140), expand = c(0,0), breaks=seq (0,200,10)) +
  scale_colour_manual(values = c("red", "#4b721d", "#cc6600", "#d3e27c", "#ce9956", "lightgreen")) +
  scale_x_discrete(limits = c("June", "July", "August", "September", "October", "November", "December", "January", "February", "March", "April", "May")) +
  theme_bw() +
  theme(strip.background = element_blank(),
        legend.background = element_blank(),
        text = element_text(size = 16, color = "black"),
        axis.text = element_text(size = 14, color = "black"),
        legend.text = element_text(size = 12, margin = margin(t = 5)),
        axis.text.x = element_text(angle = -45, vjust = 0.4, hjust = 0.1))


LFMHist_small <- ggplot(LFM_hist, aes(x = month1, y = meanLFM, color = ID)) +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM, width = 0.2), size = 1) +
  geom_line(aes(x = month1, y = meanLFM, group = ID), size = 1) +
  geom_hline(aes(yintercept = 60, color = "60% Critical Threshold"), linetype="dashed", size=1) +
  labs(x = "Month",
       y = "Live Fuel Mositure Content (%)",
       color = "Site and Species",
        title = "Average historical fuel moisture content (per month & species) \n at each collection site") +
  scale_y_continuous(limits = c(45,140), expand = c(0,0), breaks=seq (0,200,10)) +
  scale_shape_manual(values = c(15, 16, 17, 23)) +
  scale_colour_manual(values = c("red", "#4b721d", "#cc6600", "#d3e27c", "#ce9956", "lightgreen")) +
  scale_x_discrete(limits = c("June", "July", "August", "September", "October", "November", "December", "January", "February", "March", "April", "May")) +
  theme_bw() +
  theme(strip.background = element_blank(),
        legend.background = element_blank(),
        text = element_text(size = 14, color = "black"),
        axis.text = element_text(size = 10, color = "black"),
        legend.text = element_text(margin = margin(t = 5), size = 10),
        axis.text.x = element_text(angle = -45, vjust = 0.4, hjust = 0.1))

LFMHist_small
LFMHist_big

# graphing only painted cave

LFM_hist_PC <- lfm_raw %>%
  filter(concern != "Yes",
         species != "CEBE",
         species != "QUAG",
         site != "Montecito Fire",
         site != "ECC",
         site != "Tunnel Road",
         site != "St Mary's") %>%
  mutate(collection_date2 = collection_date) %>%
  separate(collection_date, sep = "-", into = c("year", "month", "day")) %>%
  mutate(month1 = month(ymd(collection_date2), label = TRUE, abbr = FALSE)) %>%
  group_by(site, species, month1) %>%
  summarise(meanLFM = mean(lfm, na.rm=TRUE),
            seLFM = std.error(lfm)) %>%
  unite(site, species, col = "ID", sep = " - ") %>%
  mutate(ID = as.factor(ID), 
    ID = fct_relevel(ID, "Painted Cave - Chamise", 
                           "St Mary's - Chamise", 
                           "Painted Cave - Bigpod ceanothus",
                           "St Mary's - Bigpod ceanothus")) %>%
  ungroup()

LFMHist_PC <- ggplot(LFM_hist_PC, aes(x = month1, y = meanLFM, color = ID)) +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM, width = 0.2), size = 1) +
  geom_line(aes(x = month1, y = meanLFM, group = ID), size = 1.25) +
  geom_hline(aes(yintercept = 60, color = "60% Critical Threshold"), linetype="dashed", size=1) +
  labs(x = "",
       y = "Live Fuel Mositure Content (%)",
       color = "Site and Species",
       title = "Average historical fuel moisture content (per month & species) \n at each collection site") +
  scale_y_continuous(limits = c(45,140), expand = c(0,0), breaks=seq (0,200,10)) +
  scale_colour_manual(values = c("red", "#4b721d", "#cc6600", "#d3e27c", "#ce9956", "lightgreen")) +
  scale_x_discrete(limits = c("June", "July", "August", "September", "October", "November", "December", "January", "February", "March", "April", "May")) +
  theme_bw() +
  theme(strip.background = element_blank(),
        legend.background = element_blank(),
        text = element_text(size = 16, color = "black"),
        axis.text = element_text(size = 14, color = "black"),
        legend.text = element_text(size = 12, margin = margin(t = 5)),
        axis.text.x = element_text(angle = -45, vjust = 0.4, hjust = 0.1))

ggsave(plot = LFMHist_big, filename = "Figures/LFM_Hist_Big.png", height = 8.5, width = 11, units = "in", dpi = 300)
ggsave(plot = LFMHist_small, filename = "Figures/LFM_Hist_Small.png", height = 4, width = 6, units = "in", dpi = 300)
ggsave(plot = LFMHist_PC, filename = "Figures/LFM_Hist_PC.png", height = 8.5, width = 11, units = "in", dpi = 300)

```

```{r 2022 web}
LFM_2022 <- lfm_raw %>%
  filter(concern != "Yes",
         species != "CEBE",
         species != "QUAG", 
         site == "Painted Cave") %>%
  mutate(collection_date2 = collection_date) %>%
  separate(collection_date2, sep = "-", into = c("year", "month", "day")) %>%
  filter(year == "2022" | year == "2023") %>%
  group_by(site, species, collection_date) %>%
  summarise(meanLFM = mean(lfm),
            seLFM = std.error(lfm)) %>%
  mutate(time = "2023",
         collection_date = as_date(collection_date))%>%
  unite(site, species, col = "ID", sep = " - ") %>%
  mutate(ID = as.factor(ID), 
    ID = fct_relevel(ID, "Painted Cave - Chamise", 
                     "Painted Cave - Bigpod ceanothus")) %>%
  ungroup()

min_2022_date <- as_date(ymd("2022-05-30"))
max_2022_date <- as_date(ymd("2023-05-30"))

LFM2022_big <- ggplot(LFM_2020, aes(x = collection_date, y = meanLFM, color = ID)) +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM), width = 7, size = 1) +
  geom_line(aes(x = collection_date, y = meanLFM, group = ID), size = 1.25) +
  geom_hline(aes(yintercept = 60, color = "60% Critical Threshold \n"), linetype="dashed", size=1) +
  labs(x = "Data Collection Date",
       y = "Live Fuel Mositure Content (%)",
       color = "Species") +
  scale_y_continuous(limits = c(45,130), expand = c(0,0), breaks=seq (0,200,10)) +
  scale_shape_manual(values = c(15, 16, 17, 23)) +
  scale_colour_manual(values = c("red", "#4b721d", "#cc6600", "#d3e27c", "#ce9956")) +
  scale_x_date(limits = c(min_2022_date, max_2022_date), breaks = "1 month", minor_breaks = NULL, date_labels = "%B", expand = c(0,0)) +
  theme_bw() +
  theme(strip.background = element_blank(),
        legend.background = element_blank(),
        text = element_text(size = 16, color = "black"),
        axis.text = element_text(size = 14, color = "black"),
        legend.text = element_text(size = 12, margin = margin(t = 5)),
        axis.text.x = element_text(angle = -45, vjust = 0.4, hjust = 0.1))

LFM2022_small <- ggplot(LFM_2020, aes(x = collection_date, y = meanLFM, color = ID)) +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM), width = 7, size = 1) +
  geom_line(aes(x = collection_date, y = meanLFM, group = ID), size = 1) +
  geom_hline(aes(yintercept = 60, color = "60% Critical Threshold \n"), linetype="dashed", size=1) +
  labs(x = "Data Collection Date",
       y = "Live Fuel Mositure Content (%)",
       color = "Species") +
  scale_y_continuous(limits = c(45,130), expand = c(0,0), breaks=seq (0,200,10)) +
  scale_shape_manual(values = c(15, 16, 17, 23)) +
  scale_colour_manual(values = c("red", "#4b721d", "#cc6600", "#d3e27c", "#ce9956")) +
  scale_x_date(limits = c(min_2022_date, max_2022_date), breaks = "1 month", minor_breaks = NULL, date_labels = "%B", expand = c(0,0)) +
  theme_bw() +
  theme(strip.background = element_blank(),
        legend.background = element_blank(),
        text = element_text(size = 14, color = "black"),
        axis.text = element_text(size = 10, color = "black"),
        legend.text = element_text(margin = margin(t = 5), size = 10),
        axis.text.x = element_text(angle = -45, vjust = 0.4, hjust = 0.1))


## change date in filename each time

#ggsave(plot = LFM2022_big, filename = "Figures/LFM_2023_11_07_Big.png", height = 8.5, width = 11, units = "in", dpi = 300)
#ggsave(plot = LFM2022_small, filename = "Figures/LFM_2023_10_07_Small.png", height = 4, width = 6, units = "in", dpi = 300)


```

