---
title: "Simple LFM Graphs"
author: "Josie Lesage"
date: "7/16/2020"
output:
  html_document:
    theme: cerulean
    code_folding: hide
    highlight: monochrome
    toc: yes
    toc_depth: 5
    toc_float: yes
    collapsed: yes
    smooth_scroll: yes
    number_sections: no
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE}

# load packages needed
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotrix)
library(reactable)
library(gridExtra)
library(lubridate)
library(here)
library(janitor)


```

# Introduction

This code is meant to produce a simple graph of LFM at sites near the Santa Barbara Botanic Garden to raise awareness of fire danger and keep local residents informed.  

Data is typically collected every two weeks from each site. Samples are collected and dried in accordance with USFS recommendations. This code has been forked and altered from Josie Lesage's original code, as data has been updated and changed.

```{r data, include=FALSE, echo=FALSE}

#read in data from csv file located in "data" folder
#use the most updated data file (indicated by date in the name of the file)
lfm_raw <- read.csv(here("data", "LFM_PaintedCave_data_072024.csv")) %>%
  clean_names() %>%  # this cleans up the data and column names
  mutate(species = recode_factor(species, #mutate lets you change column names - in this case I'm editing plant names
                                 "ADFA" = "Chamise",
                                 "CEME" = "Bigpod ceanothus")) %>%
  mutate(collection_date2 = mdy(collection_date)) %>% # changing the date column to date format 
  mutate(recode_factor(subsite, "Unknown" = "")) %>% # getting rid of "unknown" subsites
  mutate(collection_date_tidy = as.Date(collection_date, format = "%m/%d/%Y")) # putting the date into m/d/y format

lfm_raw$lfm <- as.numeric(lfm_raw$lfm)*100 #making sure lfm is a numeric character so we can perform analysis on it. If it's in a class character or anything other them numeric, then R doesn't know you can do math with it

lfm <- lfm_raw %>% # now I'm making the dataset we'll work with 
  filter(concern != "Yes",
         species != "QUAG") %>% # getting rid of any data points with "concern" or any oak points. I'll make an oak dataset later.
  group_by(collection_date, site, species, subsite) %>% #group by these subgroup options
  summarise(meanLFM = mean(lfm), # find the mean LFM, and standard deviation if there is more than one data point in a month. It takes an average of data points, and this tells you how different the data point were. THe smaller the bars, the more consistent the data is. The bigger the bars, the more the data fluctuated
            seLFM = std.error(lfm)) %>%
  mutate(site2 = site, # renaming site names. I forget why.
         species2 = species) %>%
  unite(site2, species2, col = "ID", sep = " - ") #OH! It was because we need to combine site and species into one column, but I wanted to keep the original site and species columns too
```

# Current conditions 
Here are the current conditions at Painted Cave. At some sites, only one sample is collected and there are no error bars. For sites with error bars, these represent the standard error.

```{r Graph of current conditions}
#creating data table of current conditions at PC sites
LFM_2024 <- lfm %>% #making a subset of data for 2024 (this is where I work with the most recent data we have)
  separate(collection_date, sep = "-", into = c("year", "month", "day")) %>% #I'm separating the data in the collection date column so that I can filter by year. I'm separating the column everywhere there is a dash "-" and it creates 3 columns from the one date column 
  filter(year == "2024", #only including recent data
         site == "Painted Cave", #only care about Painted Cave these days
         subsite == "PC - Lower Loop" | subsite == "PC - BP @ Oak #2") %>% 
  unite(collection_date, "year", "month", "day", sep = "-") #Reuniting collection date column now that I'm done working with it

#graphing current conditions, not including oaks        
ggplot(LFM_2024, aes(x = collection_date, y = meanLFM, color = subsite, group = (subsite))) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM, width = 0.2), size = 1) + #adding the standard deviation error bars
  geom_line(linewidth = 1) +
  labs(x = "Collection Date",
       y = "Live Fuel Mositure Content (%)",
       title = "Live Fuel Moisture Conditions") +
  scale_y_continuous(limits = c(0,190), expand = c(0,0), breaks=seq (0,200,10)) + #manually deciding the y axis tick marks
  theme_bw() + #this is a random theme that I like
  theme(axis.text.x = element_text(angle = -90, vjust = 0.25, hjust = 0.25)) +
  facet_wrap(~species)+ # I forget what this means
  scale_colour_manual(values = c("darkgreen", "#cc6600")) # choosing the colors manually

#graphing current conditions, oaks only
lfm_oak <- lfm_raw %>%
  filter(concern != "Yes",
         species == "QUAG",
         subsite != "SBBG - Tunnel") %>%
  group_by(collection_date, subsite) %>% 
  summarise(meanLFM = mean(lfm)) #no standard deviation because there's only one year of data so far, but will need this in the future

ggplot(lfm_oak, aes(x = collection_date, y = meanLFM, color = subsite)) +
  geom_point(size = 3) +
   geom_errorbar(aes(ymin = meanLFM, ymax = meanLFM, width = 0.2), size = 1) +
  geom_line(data = lfm_oak, aes(x = collection_date, y = meanLFM, color = subsite), linewidth = 1) +
  labs(x = "Collection Date",
       y = "Live Fuel Moisture Content (%)",
       title = "Live Fuel Moisture Conditions in Quercus agrifolia on Painted Cave Road") +
  scale_y_continuous(limits = c(35,175), expand = c(0,0), breaks=seq (0,175,10)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.25, hjust = 0.25)) +
  scale_colour_manual(values = c("darkred", "#8A9A5B", "#cc6600", "darkgreen", "#E8DE2A"))

reactable(
LFM_summary <- LFM_2024 %>%
  group_by(collection_date, species) %>%
  summarise(stderr = std.error(meanLFM),
            meanLFM = mean(meanLFM)) %>%
  mutate_if(is.numeric, round, digits = 2) %>%
  mutate(lfm = str_c(meanLFM, stderr, sep = " ± ")) %>%
  select(collection_date, lfm, species) %>%
  spread(collection_date, lfm)
)


```

# Historical data
We'd also like to know how current conditions compare with past data, we we'll summarize average monthly LFM for the past few years, and then will compare that line to this year's data.

```{r historical comparison}
#comparing data for all sites from 2021-2022. Hard to include 2023 because the wet year makes the error bars so high
LFM_pre_2023 <- lfm_raw %>%
  filter(concern != "Yes", 
         species != "CEBE",
         species != "QUAG", 
         site == "Painted Cave") %>%
  separate(collection_date, sep = "-", into = c("year", "month", "day")) %>%
  filter(year == "2021" | year == "2022") %>%
  group_by(site, subsite, species, month) %>%
  summarise(meanLFM = mean(lfm),
            seLFM = std.error(lfm)) %>%
  mutate(time = "mean LFM 2021-2022") %>% 
  unite(subsite, time, col = "ID", sep = " ")


#plotting 2021-2022 data (mean lfm per month)
ggplot(LFM_pre_2023, aes(x = month, y = meanLFM, color = ID, group = ID)) +
  geom_point() +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM, width = 0.2), size = 1) +
  geom_line(size = 1, aes(group = ID, color = ID)) +
  labs(x = "Collection month",
       y = "Live Fuel Mositure Content (%)",
       title = "Live Fuel Moisture Conditions at Painted Cave (average monthly value 2021-2022)") +
  scale_y_continuous(limits = c(45,140), expand = c(0,0), breaks=seq (0,200,10)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -45, vjust = -0.5, hjust = 0.5)) +
  facet_wrap(~species)


```

## For website pub.
```{r historical web}
LFM_hist <- lfm_raw %>%
  filter(concern != "Yes",
         species != "CEBE",
         species != "QUAG",
         site != "Montecito Fire",
         site != "ECC",
         site != "Tunnel Road") %>%
  mutate(collection_date2 = collection_date) %>%
  separate(collection_date, sep = "-", into = c("year", "month", "day")) %>%
  mutate(month1 = month(ymd(collection_date2), label = TRUE, abbr = FALSE)) %>%
  group_by(site, species, month1) %>%
  summarise(meanLFM = mean(lfm),
            seLFM = std.error(lfm)) %>%
  unite(site, species, col = "ID", sep = " - ") %>%
  mutate(ID = as.factor(ID), 
    ID = fct_relevel(ID, "Painted Cave - Chamise", 
                           "St Mary's - Chamise", 
                           "Painted Cave - Bigpod ceanothus",
                           "St Mary's - Bigpod ceanothus")) %>%
  ungroup()
# above plotted data from 2013-present for painted cave, but limited for St Mary's

LFMHist_big <- ggplot(LFM_hist, aes(x = month1, y = meanLFM, color = ID)) +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM, width = 0.2), size = 1) +
  geom_line(aes(x = month1, y = meanLFM, group = ID), size = 1.25) +
  geom_hline(aes(yintercept = 60, color = "60% Critical Threshold"), linetype="dashed", size=1) +
  labs(x = "",
       y = "Live Fuel Mositure Content (%)",
       color = "Site and Species",
       title = "Average historical fuel moisture content (per month & species) \n at each collection site") +
  scale_y_continuous(limits = c(45,140), expand = c(0,0), breaks=seq (0,200,10)) +
  scale_shape_manual(values = c(15, 16, 17, 23)) +
  scale_colour_manual(values = c("red", "#4b721d", "#cc6600", "#d3e27c", "#ce9956", "lightgreen")) +
  scale_x_discrete(limits = c("June", "July", "August", "September", "October", "November", "December", "January", "February", "March", "April", "May")) +
  theme_bw() +
  theme(strip.background = element_blank(),
        legend.background = element_blank(),
        text = element_text(size = 16, color = "black"),
        axis.text = element_text(size = 14, color = "black"),
        legend.text = element_text(size = 12, margin = margin(t = 5)),
        axis.text.x = element_text(angle = -45, vjust = 0.4, hjust = 0.1))


LFMHist_small <- ggplot(LFM_hist, aes(x = month1, y = meanLFM, color = ID)) +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM, width = 0.2), size = 1) +
  geom_line(aes(x = month1, y = meanLFM, group = ID), size = 1) +
  geom_hline(aes(yintercept = 60, color = "60% Critical Threshold"), linetype="dashed", size=1) +
  labs(x = "Month",
       y = "Live Fuel Mositure Content (%)",
       color = "Site and Species",
        title = "Average historical fuel moisture content (per month & species) \n at each collection site") +
  scale_y_continuous(limits = c(45,140), expand = c(0,0), breaks=seq (0,200,10)) +
  scale_shape_manual(values = c(15, 16, 17, 23)) +
  scale_colour_manual(values = c("red", "#4b721d", "#cc6600", "#d3e27c", "#ce9956", "lightgreen")) +
  scale_x_discrete(limits = c("June", "July", "August", "September", "October", "November", "December", "January", "February", "March", "April", "May")) +
  theme_bw() +
  theme(strip.background = element_blank(),
        legend.background = element_blank(),
        text = element_text(size = 14, color = "black"),
        axis.text = element_text(size = 10, color = "black"),
        legend.text = element_text(margin = margin(t = 5), size = 10),
        axis.text.x = element_text(angle = -45, vjust = 0.4, hjust = 0.1))

LFMHist_small
LFMHist_big

ggsave(plot = LFMHist_big, filename = "Figures/LFM_Hist_Big.png", height = 8.5, width = 11, units = "in", dpi = 300)
ggsave(plot = LFMHist_small, filename = "Figures/LFM_Hist_Small.png", height = 4, width = 6, units = "in", dpi = 300)

```

```{r 2022 web}
LFM_2022 <- lfm_raw %>%
  filter(concern != "Yes",
         species != "CEBE",
         species != "QUAG", 
         site == "Painted Cave") %>%
  mutate(collection_date2 = collection_date) %>%
  separate(collection_date2, sep = "-", into = c("year", "month", "day")) %>%
  filter(year == "2022" | year == "2023") %>%
  group_by(site, species, collection_date) %>%
  summarise(meanLFM = mean(lfm),
            seLFM = std.error(lfm)) %>%
  mutate(time = "2022",
         collection_date = as_date(mdy(collection_date)))%>%
  unite(site, species, col = "ID", sep = " - ") %>%
  mutate(ID = as.factor(ID), 
    ID = fct_relevel(ID, "Painted Cave - Chamise", 
                     "Painted Cave - Bigpod ceanothus")) %>%
  ungroup()

min_2022_date <- as_date(ymd("2022-05-30"))
max_2022_date <- as_date(ymd("2023-05-30"))

LFM2022_big <- ggplot(LFM_2022, aes(x = collection_date, y = meanLFM, color = ID)) +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM), width = 7, size = 1) +
  geom_line(aes(x = collection_date, y = meanLFM, group = ID), size = 1.25) +
  geom_hline(aes(yintercept = 60, color = "60% Critical Threshold \n"), linetype="dashed", size=1) +
  labs(x = "Data Collection Date",
       y = "Live Fuel Mositure Content (%)",
       color = "Species") +
  scale_y_continuous(limits = c(45,130), expand = c(0,0), breaks=seq (0,200,10)) +
  scale_shape_manual(values = c(15, 16, 17, 23)) +
  scale_colour_manual(values = c("red", "#4b721d", "#cc6600", "#d3e27c", "#ce9956")) +
 scale_x_date(breaks = "1 month", minor_breaks = "1 week", labels=date_format("%B")) +
  theme_bw() +
  theme(strip.background = element_blank(),
        legend.background = element_blank(),
        text = element_text(size = 16, color = "black"),
        axis.text = element_text(size = 14, color = "black"),
        legend.text = element_text(size = 12, margin = margin(t = 5)),
        axis.text.x = element_text(angle = -45, vjust = 0.4, hjust = 0.1))

LFM2022_small <- ggplot(LFM_2020, aes(x = collection_date, y = meanLFM, color = ID)) +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM), width = 7, size = 1) +
  geom_line(aes(x = collection_date, y = meanLFM, group = ID), size = 1) +
  geom_hline(aes(yintercept = 60, color = "60% Critical Threshold \n"), linetype="dashed", size=1) +
  labs(x = "Data Collection Date",
       y = "Live Fuel Mositure Content (%)",
       color = "Species") +
  scale_y_continuous(limits = c(45,130), expand = c(0,0), breaks=seq (0,200,10)) +
  scale_shape_manual(values = c(15, 16, 17, 23)) +
  scale_colour_manual(values = c("red", "#4b721d", "#cc6600", "#d3e27c", "#ce9956")) +
  scale_x_date(limits = c(min_2022_date, max_2022_date), breaks = "1 month", minor_breaks = NULL, date_labels = "%B", expand = c(0,0)) +
  theme_bw() +
  theme(strip.background = element_blank(),
        legend.background = element_blank(),
        text = element_text(size = 14, color = "black"),
        axis.text = element_text(size = 10, color = "black"),
        legend.text = element_text(margin = margin(t = 5), size = 10),
        axis.text.x = element_text(angle = -45, vjust = 0.4, hjust = 0.1))


## change date in filename each time

ggsave(plot = LFM2022_big, filename = "Figures/LFM_2023_11_07_Big.png", height = 8.5, width = 11, units = "in", dpi = 300)
ggsave(plot = LFM2022_small, filename = "Figures/LFM_2023_10_07_Small.png", height = 4, width = 6, units = "in", dpi = 300)


```

