---
title: "RWMP Summary"
author: "Caroline Shepherd"
date: "2024-04-23"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotrix)
library(reactable)
library(gridExtra)
library(lubridate)
library(here)
library(janitor)

# packages installed
```

# Introduction

This code is meant to produce a simple graph of LFM at sites near the Santa Barbara Botanic Garden to raise awareness of fire danger and keep local residents informed.  

Data is typically collected every two weeks from each site. Samples are collected and dried in accordance with USFS recommendations. This code has been forked and altered from Josie Lesage's original code, as data has been updated and changed.

```{r data, include=FALSE, echo=FALSE}

#reading in data, cleaning it
rwmp_raw <- read.csv(here("data", "LFM_PaintedCave_data_042024.csv")) %>%
  clean_names() %>% 
  mutate(species = recode_factor(species,
                                 "ADFA" = "Chamise",
                                 "CEME" = "Bigpod ceanothus")) %>%
  mutate(recode_factor(subsite, "Unknown" = ""))

#changing "collection date" column to date class in new column, "date"
rwmp_raw$date <- parse_date_time(rwmp_raw$collection_date, "ymd")

#checking to see if lubridate worked (it did)
class(rwmp_raw$date)

#checking class of lfm bc it needs to be number for analysis
class(rwmp_raw$lfm_perc)
class(rwmp_raw$lfm)

#changing class
rwmp_raw$lfm_perc <- as.numeric(rwmp_raw$lfm_perc)

#check to see if it worked 
class(rwmp_raw$lfm_perc)
#it did! class is now "numeric"

#creating new filtered dataframe with mean LFM values and new site column
lfm <- rwmp_raw %>%
  filter(concern != "Yes") %>%
  group_by(date, site, species, subsite) %>%
  summarise(meanLFM = mean(lfm_perc),
            seLFM = std.error(lfm_perc)) %>%
  mutate(site2 = site,
         species2 = species) %>%
  unite(site2, species2, col = "ID", sep = " - ")

```

# Current conditions 
Here are the current conditions along Painted Cave Road

```{r Graph of current conditions}
#creating data table of current conditions at PC sites
lfm2023 <- lfm %>%
  separate(date, sep = "-", into = c("year", "month", "day")) %>%
  filter(year == "2023" | year == "2024",
         site == "Painted Cave",
         subsite == "PC - Lower Loop" | subsite == "PC - BP @ Oak #2") %>% 
  unite(collection_date, "year", "month", "day", sep = "-")

lfm2023 <- lfm %>% 
  filter(year == "2023",
         year == "2024")

#graphing current conditions, not including oaks        
ggplot(lfm2023, aes(x = collection_date, y = meanLFM, color = subsite, group = (subsite))) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = meanLFM-seLFM, ymax = meanLFM+seLFM, width = 0.2), size = 1) +
  geom_line(linewidth = 1) +
  labs(x = "Collection Date",
       y = "Live Fuel Mositure Content (%)",
       title = "Live Fuel Moisture Conditions") +
  scale_y_continuous(limits = c(0,190), expand = c(0,0), breaks=seq (0,200,10)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.25, hjust = 0.25)) +
  facet_wrap(~species)+
  scale_colour_manual(values = c("darkgreen", "#cc6600"))

#graphing current conditions, oaks only
lfm_oak <- lfm %>%
  separate(date, sep = "-", into = c("year", "month", "day")) %>%
  filter(species == "QUAG",
         subsite != "SBBG - Tunnel") %>%
  unite(collection_date, "year", "month", "day", sep = "-") %>% 
  group_by(collection_date, subsite)
#no standard deviation because there's only one year of data so far, but will need this in the future

ggplot(lfm_oak, aes(x = collection_date, y = meanLFM, color = subsite)) +
  geom_point(size = 3) +
   geom_errorbar(aes(ymin = meanLFM, ymax = meanLFM, width = 0.2), size = 1) +
  geom_line(data = lfm_oak, aes(x = collection_date, y = meanLFM, color = subsite), linewidth = 1) +
  labs(x = "Collection Date",
       y = "Live Fuel Moisture Content (%)",
       title = "Live Fuel Moisture Conditions in Quercus agrifolia on Painted Cave Road") +
  scale_y_continuous(limits = c(35,175), expand = c(0,0), breaks=seq (0,175,10)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = -90, vjust = 0.25, hjust = 0.25)) +
  scale_colour_manual(values = c("darkred", "#8A9A5B", "#cc6600", "darkgreen", "#E8DE2A"))

class(lfm_oak$meanLFM)

reactable(
LFM_summary <- LFM_2023 %>%
  group_by(collection_date, species) %>%
  summarise(stderr = std.error(meanLFM),
            meanLFM = mean(meanLFM)) %>%
  mutate_if(is.numeric, round, digits = 2) %>%
  mutate(lfm = str_c(meanLFM, stderr, sep = " ± ")) %>%
  select(collection_date, lfm, species) %>%
  spread(collection_date, lfm)
)


```